---
title: "Worksheet9"
author: "STAT414"
date: "2024-11-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Module09 deals with two important topics: obtaining prediction intervals for future observations from a process, and determining tolerance intervals that are expected to capture certain proportion of individuals from the population. These concepts are closely related to the concept of quantiles introduced earlier. Prediction intervals are used widely in many applications, especially in the area of statistical quality control. Tolerance intervals are used in biomedical fields, e specially in determining acceptable ranges of biomarkers (such as level of sugar in blood samples, typical range of cholesterol levels in certain populations etc).

```{r}
#1. Suppose a population is well approximated by a Normal distribution with 
#mu = 10 and sig = 2. 
#2. Suppose a population is well-modeled by a lognormal distribution with theta = 4 
#and tau = 1. 

mu = 10
sig = 2
theta = 4
tau = 1

#(a) Obtain the quantiles for this normal population for p = 0.025, 0.05, 0.95, 0.975 

p = c(0.025, 0.05, 0.95, 0.975)

quantiles <- data.frame (
  population = "N(mu=10, sig=2)",
  quantiles = p,
  values = qnorm(p, mean = mu, sd = sig)
)

#(a) Obtain the quantiles for this lognormal population for p = 0.025, 0.05, 0.95, 0.975 

cat("A percentile is a quantile corresponding to p=[0,1], with Xp corresponding to 
the .p quantiles. For any continuous distribution, the CDF F can be used to calculate
a quantile: Xp = F^-1(p)")

append <- data.frame (
  population = "LN(theta=4, tau=1)",
  quantiles = p,
  values = qlnorm(p, meanlog = theta, sdlog = tau)
)

#combine both populations and the calculated quantiles into 1 df
quantiles <- rbind(quantiles, append)
quantiles

#(b) Determine a tolerance interval with 90% coverage. Determine a TI with a 
#95% coverage. 

cat("A tolerance interval for a population contains the pre-specified proportion
    of the population. A Beta * 100% TI is called the coverage of the interval.
    Create 2 samples using the normal and lognormal parameters, since there is
    no actual dataset provided, and a vector x is a required input for the EnvStats
    function to generate TI's tolIntNorm and tolIntLnorm. Choose a high sample size
    since we are trying to calculate theoretical values.
    Choose alpha=.05")

# norm_sample <- rnorm(10000, mean=mu, sd=sig)
# lnorm_sample <- rlnorm(10000, meanlog=theta, sdlog=tau)
# 
# #Normal distribution 90% TI & 95% TI
# tolIntNorm(norm_sample, coverage = 0.90, ti.type = "two-sided", conf.level = 0.95, 
#            method = "exact")
# tolIntNorm(norm_sample, coverage = 0.95, ti.type = "two-sided", conf.level = 0.95, 
#            method = "exact")
# 
# #Lognormal distribution 90% TI & 95% TI
# tolIntLnorm(lnorm_sample, coverage = 0.90, ti.type = "two-sided", conf.level = 0.95, 
#            method = "exact")
# tolIntLnorm(lnorm_sample, coverage = 0.95, ti.type = "two-sided", conf.level = 0.95, 
#            method = "exact")

cat("There is a simpler method to calculate tolerance intervals using quantiles,
    by simply using quantiles containing the desired proportions. Using .5 +- Beta/2
    as the quantiles of the lower and upper bounds for the TI.")

lower_90 <- .5 - (.9/2)
upper_90 <- .5 + (.9/2)
lower_95 <- .5 - (.95/2)
upper_95 <- .5 + (.95/2)

cat(sprintf("90%% Normal TI: (%.2f, %.2f)\n", 
            qnorm(lower_90, mean = mu, sd = sig), 
            qnorm(upper_90, mean = mu, sd = sig)))
cat(sprintf("95%% Normal TI: (%.2f, %.2f)\n", 
            qnorm(lower_95, mean = mu, sd = sig), 
            qnorm(upper_95, mean = mu, sd = sig)))
cat(sprintf("90%% Lognormal TI: (%.2f, %.2f)\n", 
            qlnorm(lower_90, meanlog = theta, sdlog = tau), 
            qlnorm(upper_90, meanlog = theta, sdlog = tau)))
cat(sprintf("95%% Lognormal TI: (%.2f, %.2f)\n",
            qlnorm(lower_95, meanlog = theta, sdlog = tau), 
            qlnorm(upper_95, meanlog = theta, sdlog = tau)))

#(c) For each coverage level, provide several other intervals of the form 
#[xdelta, xbeta+delta] with the same coverage by taking different values of delta. 

#For any delta < 1-beta, [Xdelta, Xdelta+beta] is a B 100% TI
#delta + beta < 1

#for b=90, deltamin = .01, deltamax = .1
delta_90 <- seq(.01, .1, by=.01)
delta_90 <- delta_90[delta_90 + 0.90 < 1]

#for b=95, deltamin = .01, deltamax = .05
delta_95 <- seq(.01, .05, by=.01)
delta_95 <- delta_95[delta_95 + 0.95 < 1] 
  
for(i in delta_90){
  cat(sprintf("Delta: %.2f | 90%% Normal TI: (%.2f, %.2f)\n", 
              i,
              qnorm(i, mean = mu, sd = sig), 
              qnorm(i + .90, mean = mu, sd = sig)))

  cat(sprintf("Delta: %.2f | 90%% Lognormal TI: (%.2f, %.2f)\n", 
              i,
              qlnorm(i, meanlog = theta, sdlog = tau), 
              qlnorm(i + .90, meanlog = theta, sdlog = tau)))
}

for(i in delta_95){
  cat(sprintf("Delta: %.2f | 95%% Normal TI: (%.2f, %.2f)\n",
              i,
              qnorm(i, mean = mu, sd = sig), 
              qnorm(i + .95, mean = mu, sd = sig)))

  cat(sprintf("Delta: %.2f | 95%% Lognormal TI: (%.2f, %.2f)\n", 
              i,
              qlnorm(i, meanlog = theta, sdlog = tau), 
              qlnorm(i + .95, meanlog = theta, sdlog = tau)))
}


#(d) Plot the length of these intervals vs delta. Comment. 

  #create dataframes to store the lengths of the TIs for different deltas
  delta_90_lengths <- data.frame(delta = delta_90)
  delta_95_lengths <- data.frame(delta = delta_95)
  
  #For each coverage value, calculate the LB and UB for normal and lognormal
  #populations, for each value of delta for the corresponding tolerance levels,
  #and store in the corresponding dataframes
  delta_90_lengths$normal_lower <- qnorm(delta_90, mean = mu, sd = sig) 
  delta_90_lengths$normal_upper <- qnorm(delta_90 + .90, mean = mu, sd = sig)
  
  delta_90_lengths$lnormal_lower <- qlnorm(delta_90, meanlog = theta, sdlog = tau)
  delta_90_lengths$lnormal_upper <- qlnorm(delta_90 + .90, meanlog = theta, sdlog = tau)
  
  delta_95_lengths$normal_lower <- qnorm(delta_95, mean = mu, sd = sig)
  delta_95_lengths$normal_upper <- qnorm(delta_95 + .95, mean = mu, sd = sig)
  
  delta_95_lengths$lnormal_lower <- qlnorm(delta_95, meanlog = theta, sdlog = tau)
  delta_95_lengths$lnormal_upper <- qlnorm(delta_95 + .95, meanlog = theta, sdlog = tau)
  
  #Now for each normal and lognormal population, calculate the lengths of the found
  #tolerance intervals for B=.9 and B=.95
  delta_90_lengths$normal_length <- delta_90_lengths$normal_upper- delta_90_lengths$normal_lower
  delta_90_lengths$lnormal_length <- delta_90_lengths$lnormal_upper- delta_90_lengths$lnormal_lower

  delta_95_lengths$normal_length <- delta_95_lengths$normal_upper- delta_95_lengths$normal_lower
  delta_95_lengths$lnormal_length <- delta_95_lengths$lnormal_upper- delta_95_lengths$lnormal_lower
  
  # #print lengths
  delta_90_lengths$normal_length
  delta_90_lengths$lnormal_length

  delta_95_lengths$normal_length
  delta_95_lengths$lnormal_length

# Plot lengths for normal population
plot(delta_90_lengths$delta, delta_90_lengths$normal_length, type = "b",
  col = "blue", ylim = c(4,max(delta_95_lengths$normal_length)+4),
  xlim = c(.01,.1), xlab = "Delta", ylab = "Interval Length", 
  main = "Normal Tolerance Interval Lengths")
lines(delta_95_lengths$delta, delta_95_lengths$normal_length, type = "b", col = "red")
legend("topright", legend = c("90% Coverage", "95% Coverage"), col = c("blue", "red"),
       lty = 1)

# Plot lengths for lognormal population
plot(delta_90_lengths$delta, delta_90_lengths$lnormal_length, type = "b",
  col = "blue", ylim = c(100,max(delta_95_lengths$lnormal_length)+100),
  xlim = c(.01,.1), xlab = "Delta", ylab = "Interval Length", 
  main = "Lognormal Tolerance Interval Lengths")
lines(delta_95_lengths$delta, delta_95_lengths$lnormal_length, type = "b", col = "red")
legend("bottomright", legend = c("90% Coverage", "95% Coverage"), col = c("blue", "red"),
       lty = 1)

cat("From the plots we can see that as delta increases, the lengths of the intervals
    increase for lognormal distributions; there is a positive correlation. However for
    a normal distribution the delta affects the interval length parabolically, meaning that
    extreme values being the min/max of delta are where the TI interval lengths are 
    maximized.")

#(e) Obtain a random sample of size n = 25 from this population. Obtain 90% TI 
#from your sample. Compare the TI based on exact TI obtained in part (a).

library(EnvStats)
set.seed(1)

#n=25 samples for normal and lognormal populations
norm_sample <- rnorm(25, mean=mu, sd=sig)
lnorm_sample <- rlnorm(25, meanlog=theta, sdlog=tau)

#Normal distribution 90% TI
sample_norm <- tolIntNorm(norm_sample, coverage = 0.90, ti.type = "two-sided", conf.level = 0.95,
           method = "exact")
sample_norm

#Lognormal distribution 90% TI
sample_lnorm <- tolIntLnorm(lnorm_sample, coverage = 0.90, ti.type = "two-sided", conf.level = 0.95,
           method = "exact")
sample_lnorm


cat(sprintf("n=25 Normal sample 90%% TI: (%.2f, %.2f)\n",
            sample_norm$interval$limits[1], sample_norm$interval$limits[2]))
cat(sprintf("90%% Normal TI: (%.2f, %.2f)\n", 
            qnorm(lower_90, mean = mu, sd = sig), 
            qnorm(upper_90, mean = mu, sd = sig)))

cat(sprintf("n=25 Lognormal sample 90%% TI: (%.2f, %.2f)\n",
            sample_lnorm$interval$limits[1], sample_lnorm$interval$limits[2]))
cat(sprintf("90%% Lognormal TI: (%.2f, %.2f)\n", 
            qlnorm(lower_90, meanlog = theta, sdlog = tau), 
            qlnorm(upper_90, meanlog = theta, sdlog = tau)))

cat("For normal and lognormal populations, it is expected that the sample calculation
    of the tolerance intervals be different than the actual tolerance intervals, when
    using a sample of n=25. This is because the sample size is small, and is less
    likely to accurately represent the population parameters, especially
    for lognormal since it is inherently skewed. From the output above we can see
    that the normal sample calculation does differ slighty from the actual
    tolerance interval, and the lognormal sample calculation is significantly different
    from the actual TI.")
```