---
title: "Untitled"
output:
  html_document:
    df_print: paged
date: "2023-06-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a function to compute the statistic. Here we are using the example 
$T= \frac{X_1+X_2}{2}$. You can change this part of the program to any other
function as you need. 

```{r}
fun <- function(x) {
mean(x)
}
```


Generating a large number of samples, each of the sample size as specified. 
The default simulation size $(M)$ used here is 1000, and the default 
sample size $(n)$ used here is 2. 

The Pop vector below should contain the entire list of population values. Repeated 
values (that is, two differnet population units having the same meausurements) should be 
all be listed. We are using the example on pages 5-7 of the Module 04 slides. Hence:

Pop <- c(12,7,5)  
sampsize <- 2


```{r}
simsize <- 1000
sampsize <- 2
set.seed(2)
Pop <- c(12,7,5)
Samples <- as.data.frame(matrix(sample(Pop, simsize*sampsize, replace=TRUE),ncol=sampsize))
rownames(Samples) <- paste("sample", 1:simsize, sep="")
colnames(Samples) <- paste("obs", 1:sampsize, sep="")
#Samples
```


Compute the statistic for each of the generated simsize samples. We are using a useful
R function called apply, which can implement the same function for each row. 
This is a handy tool when we have to do the same operation on each row for a very large
matrix of data. 

```{r}
results <- apply(Samples,FUN=fun,1)
#results
```


Now, we plot the histogram of all simulated values of the test statistic $T$, and compare it to 
a Normal distribution, with $\mu$ and $\sigma$ approximated by the average and standad deviation
of all simulated values of T}$. 


```{r}
hist(results,prob=T,breaks=4.5:12.5,main="",xlab="xbar")
#curve(dnorm(x,mean=mean(results),sd=sd(results)), from = min(results), to = max(results),add=T,col="red") 
curve(dnorm(x,mean=mean(results),sd=sd(results)), from = 4.5, to = 12.5,add=T,col="red") 
title(main="Simulated Sampling Distribution")
```


Let us now compare this to the exact computation. Recall, from the M04 slides, the sampling distribution of T is given by

\begin{center}
\begin{tabular}{|c|r|r|r|r|r|r|}
\hline
t & 5 & 6 & 7 & 8.5 & 9.5 & 12 \\
\hline
$P(T=t)$ & $\frac{1}{9}$ & $\frac{2}{9}$ & $\frac{1}{9}$ & $\frac{2}{9}$ & $\frac{2}{9}$ & $\frac{1}{9} $ \\ 
\hline    
\end{tabular}    
\end{center}

```{r}
xbar <- c(5,6,6,7,8.5,8.5,9.5,9.5,12)
#barplot(table(xbar))
hist(xbar,prob=T,breaks=4.5:12.5,ylim=c(0,0.25),main="")
title(main="Exact Sampling Distribution")
```


