---
title: 'Module04: Sampling Distributions'
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2023-06-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Create a function to compute the statistic. Here we are using the example 
$T= \frac{X_1+X_2}{2}$. You can change this part of the program to any other
function as you need. 

```{r}
fun <- function(x) {
mean(x)
}
```


Generating a large number of samples, each of the sample size as specified. 
The default simulation size $(M)$ used here is 1000, and the default 
sample size $(n)$ used here is 2. 

The population is assumed to a Normal distribution with $\mu=20$ and $\sigma=6$. 
We will consider a sample size of $n=2$. 
That is, 

sampsize <- 2


```{r}
simsize <- 10000
sampsize <- 4
set.seed(2)
Samples <- as.data.frame(matrix(rnorm(simsize*sampsize, mean=10,sd=2),ncol=sampsize))
rownames(Samples) <- paste("sample", 1:simsize, sep="")
colnames(Samples) <- paste("obs", 1:sampsize, sep="")
#Samples
```


Compute the statistic for each of the generated simsize samples. We are using a useful
R function called apply, which can implement the same function for each row. 
This is a handy tool when we have to do the same operation on each row for a very large
matrix of data. 

```{r}
results <- apply(Samples,FUN=fun,1)
Samples
#results
```


Now, we plot the histogram of all simulated values of the test statistic $T$, and compare it to 
a Normal distribution, with $\mu$ and $\sigma$ approximated by the average and standad deviation
of all simulated values of T}$. 


```{r}
hist(results,prob=T)
curve(dnorm(x,mean=mean(results),sd=sd(results)), from = min(results), to = max(results),add=T,col="red") 
#curve(dnorm(x,mean=mean(results),sd=sd(results)), from = 4.5, to = 12.5,add=T,col="red") 
```


