
---
title: "Bootstrapping to Compute SE of a statistic, Example 2"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
date: "2023-16-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## A generic code for bootstrapping to estimate the standard error of a statistic
#   Example uses RExam data set and illustrates the Bootstrap estimate for the standard error 
#   of the sample first quartile (Q1)
#   RExam data is analysized in DSUR

```{r}
Rexam <-
RExam <- 
  read.delim("C:/Users/criss/Desktop/STAT414/Assignments/10-19/RExam.dat")
```

Plotting the origina data: From RExam, we are going to consider the computer scores. 
The objective is to estimate the population quratile (Q1) for this population 
and estimate its standard error by bootstrapping.
 
## Validity of bootstrapping requires only an assumption that 
## *  the data is collected according to a SRSWR design
## No specific distributional asusmption is made about the collected data


Step 1. Estimate of the population quartile by using the sample quartile:

```{r}
with(Rexam, hist(computer, prob=T, breaks="Sturges",
col="darkgray"))
sampQ1 <- quantile(Rexam$computer,0.25)
```

Step 2. Perform bootstrapping  (draw samples from the observed data using the same sampling design 
as the one used to obtain the original data)
 
# Generate 10000 Bootstrap samples of each size n=100 from Rexam$computer 
# Resulting BS samples are save in a matrix called results


```{r}
BS.Samples <- as.data.frame(matrix(sample(Rexam$computer,size=10000*100, replace=TRUE),ncol=100))
rownames(BS.Samples) <- paste("BSsample", 1:10000, sep="")
colnames(BS.Samples) <- paste("obs", 1:100, sep="")
```

# Create a function to compute the sample quantile. 

  You can use any built in R function or build your own. 


```{r}
funquant <- function(x) {
quantile(x,c(0.25))
}
```

# Apply the function to each row of 

# the ensamble of bootstrap samples. 

```{r}
results <- apply(BS.Samples,FUN=funquant,1)
```


# Compute summary statistics

Step 3. Obtain bootstap estimates of the SE and the sampling distribution of sampQ1

  The standard deviation reported here is the bootstrap standard error of estimated quartile (sampQ1)
  The histogram reported below is the bootstrap estimate of sampling distribution of sample quartile

```{r}
summary(results)
sd(results)
hist(results,prob=T,breaks=8)
```


